### 2024-04-07 thesis/input_files/analysis.tex
April 7 2024

>>>>

\section{Point forecasting}
This section carries out a comparative study between the state of the art methods for point forecasting, introduced in \ref{ch:point}.
As use case, we will consider the task of predicting the load for the next month.
In such setting we have the following regressors
\begin{itemize}
    \item Day
    \item Hour
    \item Month
    \item Day of the week
    \item Is holiday
    \item Weather temperature
\end{itemize}
The metric upon we agreed is the root mean squared error, see section \ref{metrics}.
\subsection{Multiple linear regression}
To get started, standard multiple linear regression has been applied, see fig \ref{fig:mlr_price} for a visualisation. 
\begin{figure}
    \includegraphics[width=\textwidth]{images/mlr_price.png}
    \caption{Multiple linear regression prediction}
    \label{fig:mlr_price}
\end{figure}

The resulting RMSE is 30.59.
What can be concluded, is that multiple linear regression is capable of catching the daily seasonality. Nevertheless, it cannot catch the range of the price series properly.

\subsection{Tbats}
Next, we tried with the autoregressive approach. Unfortunately, AR, ARIMA and SARIMA models did not perform as expected, their output was slightly better than the one of linear regression. This is probably due to the fact that, the data considered entails two kinds of seasonalities, while ARIMA models can only handle one at a time. We remind the reader, that the price time series involves both daily and weekly seasonalities. Hence, the need for a more advanced time series model. The Tbats model is a forecasting method capable of handling complex patterns in the data. Its name stands for trigonometric seasonality, Box-Cox transformation, ARMA errors, trend and seasonal components. 
Tbats forecast is visualized in figure \ref{fig:tbats_price}, meanwhile it RMSE is 15.08.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/tbats_price.png}
    \caption{Tbats prediction}
    \label{fig:tbats_price}
\end{figure}
From the plot we see that Tbats is capable of catching the lows and average trend, conversely it has some difficulties handling the price peaks.
\subsection{Prophet}
Following, the prophet model has been considered. 
We get started by considering the base implementation. In such setting, prophet takes in as input the time series object and learns its data generating process.
This method achieves a RMSE of 23.96, its prediction is visualized in figure \ref{fig:prophet_price_1}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price_1.png}
    \caption{Prophet prediction}
    \label{fig:prophet_price_1}
\end{figure}
What can be seen is that, prophet models correctly the average trend but does not model peaks and lows precisely. 
Next, a more complex model was trained. We added the weather temperature, the square of it and the categorical variable for holidays effect as regressors. Furthermore, we also applied a log transformation to the dependet variable. Doing so, RMSE went down to 10.29. Forecast is visualized in \ref{fig:prophet_price2}, moreover, figure \ref{fig:prophet_price2.1} and figure \ref{fig:prophet_price2.2} break down the trend and the seasonalities respectively.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.png}
    \caption{Prophet v2 prediction}
    \label{fig:prophet_price2}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.1.png}
    \caption{Seasonalities breakdown}
    \label{fig:prophet_price2.1}
\end{figure}

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/prophet_price2.2.png}
    \caption{Prices trend}
    \label{fig:prophet_price2.2}
\end{figure}
Prophet alongside with the right features is a good model in the context of electricity price forecasting, it is capcable of catching the trend, the peak and the lows.

\subsection{K-nearest neighbours}
Afterwards, K-nearest neighbours has been considered. In doing so, we cross validated the number of neighbours to get the best model instance. The forecast is visualized in figure \ref{fig:knn_price}, RMSE is 12.54.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/knn_price.png}
    \caption{K-nearest neighbour regression}
    \label{fig:knn_price}
\end{figure}
What it can be said is that, K-nearest neighbour is capable of predicting prices well by averaging past data.

\subsection{Support vector regression}
Coming up we have support vector regression. In applying this model we used gridsearch crossvalidation to search for the best regularization parameter C. Support vector regression achieves an RMSE of 23.24, for the prediction see figure \ref{fig:svr_price}.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/svr_price.png}
    \caption{Support vector regression prediction}
    \label{fig:svr_price}
\end{figure}
Visually, we see that the SVR performs similar to multiple linear regression as expected. Similarly to multiple linear regression, SVR can model the daily seasonality but the point prediction is off in terms of highs and lows.

\subsection{LSTM}

\subsection{Kernel ridge regression}
Subsequently, kernel ridge regression was considered. The kernel considered is the radial basis gaussian function.
Cross validation was carried jointly over the rbf kernel bandwith and the regularization constant.
The RMSE achieved is 9.86, figure \ref{fig:krnridge_price} reports the prediction.
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/krnridge_price.png}
    \caption{Kernel ridge prediction}
    \label{fig:krnridge_price}
\end{figure}
We can observe that, kernel ridge regression models accurately the electricity time series.

\subsection{Kernel support vector regression}
The last model we considered in this setting was the kernel support vector regression.
RMSE achieved with this method is 9.10, forecast is reported in figure \ref{fig:krnsvr_price}.

\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/krnsvr_price.png}
    \caption{Kernel support vector prediction}
    \label{fig:krnsvr_price}
\end{figure}
We can observe that kernel support vector prediction is one of the best performing techniques between the one considered.

\subsection{Results}
This section reports the table comparing the considered methods in the context of electricity load forecasting.

- table 
|methods|tasks RMSE|
|       |X          |       
|       |X          |

### 2024-04-07 thesis/input_files/data_exploratory.tex
April 7 2024

>>>>

This section covers the datasets used in our experiments, attributes and features will be thoroughly described.
Next, we will explain the ETL pipeline that we set up in order to ease the workflow of our comparison studies.
Finally, in order to to bettern understand patters within the data, we carry out an exploratory data analysis.

\section{Dataset}
The chosen datasets for probabilistic load forecasting, were the data from the GEFCom 2014 competition. The data is freely hosted on Dr. Hong blog \cite{hong2016probabilistic}.
The main reason was that, these dataset are considered an a excellent test case for comparing predictive models between the EF community. Additionally, the scores of the competing models are freely available, this enables us to carry out a clean and transparent comparative study.
The GEFCom 2014 consisted of four tracks: price, load, wind power and solar power forecasting. 
The GEFCom2014 folder contains a zip file for each of the four tracks.
This thesis work is focuesed on providing forecasts for the load and price quantities.
\\
\subsection{Price track}
In the price forecasting track, the goal was to predict the electricity price for the next 24 hours of a single zone. 
The data can be found in the GEFCom2014-P\_V2 zip file alongside a set of instructions and the benchmark forecasts.
This data consisted of time series for the locational marginal price, for the zonal load and for the system load. The data covers the time interval ranging from the 1st of January 2011 to the 15th of June 2013, see figure \ref{fig:price_track_fig1}. 
Additionally, as the competition went on, the real observed data of the previous tasks were made available.
The fifteen target days ordered by task number are: 16/06/2013, 17/06/2013, 24/06/2013, 04/07/2013, 09/07/2013, 13/07/2013, 16/07/2013, 18/07/2013, 19/07/2013, 20/07/2013, 24/07/2013, 25/07/2013, 07/12/2013, 08/12/2013. 
\begin{figure}[!h]
    \includegraphics[width=\textwidth]{images/price_track_fig1.png}
    \caption{Price track}
    \label{fig:price_track_fig1}
\end{figure}
\subsection{Load track}
In the load track, contestant were asked to provide one month ahead hourly probabilistic forecasts on a rolling basis for 15 consecutive rounds. To get started, in the first round, organizers provided 69 months of hourly load data (from 01/01/2005 to 30/10/2010) and 117 months of hourly temperature data (from 01/01/2001 to 30/10/2010). 
As for the price track, the true observed data from the previous track were made available as the competition progressed.
\\
The data for load forecasting is contained in the GEFCom2014-L\_V2 zip file. Within this subfolder, we a txt with the competition instructions and 15 folders one for each of the consecutive tasks. For each of those we have the prediction of the competition benchmark model and the train file upon which we will fit our model.
Each data from the various task folders are shifted by one months between others.
\\
In order to compare our model performance with the winning entries of the GEFCom2014 load competition, we will refer to the Provisional\_Leaderboard\_V2 file contained also in the GEFCom2014 Data directory.
% Notice a quick premise, within this xsls file the authors opted for various scores to compare the entries performance. This choice was motivated by the need 
% to give a proportial weight to each task when averaging between them; that is in order to not penalise too much the tasks with higher point spreads. Furthermore, additional factors such as frequency of benchmark outperformance, and robustness and clarity in methodology explaination.
\\
Notice, the pinball scores for the load track are stored inside the subtab L-score-0/L-score-2 of Provisional\_Leaderboard\_V2.

\section{ETL}

\section{EDA}

### 2024-04-07 thesis/input_files/implementation.tex
April 7 2024

- As I explain code scripts go over the test, to 
explain better my ideas.
\section{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\section{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\section{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\section{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
\subsection{Python vs R implementation}
In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.

>>>>

- As I explain code scripts go over the test, to 
explain better my ideas.

- Indicate also hyperparameters maybe in each subsection

\section{Point forecasting}
\subsection{Multiple linear regression}
For multiple linear regression the one from the \href{https://scikit-learn.org/stable/}{sklearn} library has been used.

\subsection{Tbats}
Tbats implementation is available at \url{https://github.com/intive-DataScience/tbatsx}.
In our application we specified as hyperparameters the lenght of seasons; 24 for the daily seasonality and 168 for the weekly seasonality.

\subsection{Prophet}
The prophet models has been applied by employing the python API provided by Meta \href{https://facebook.github.io/prophet/docs/quick_start.html}{prophet}.

\subsection{K-nearest neighbours}
The object KNeighborsRegressor of the sklearn module neighbors has been used.

\subsection{Support vector regression}
The object SVR of the sklearn module svm has been used by specifying the liner kernel.

\subsection{LSTM}

\subsection{Kernel ridge regression}
The object KernelRidge of the sklearn module kernel\_ridge has been used.

\subsection{Kernel support vector regression}
The object SVR of the sklearn module svm has been used by specifying rbf as the kernel parameter.

\section{Probabilistic forecasting}
\subsection{Linear quantile regression}
The implementation of statsmodels.regression.quantile\_regression.QuantReg has been used.
The model is fitted through iterative reweighted least squares.
\subsection{Quantile gradient boosting machine}
The implementation of sklearn.ensemble.GradientBoostingRegressor has been used.
\subsection{Quantile forest}
The implementation of quantile\_forest.RandomForestQuantileRegressor has been used. This estimator is compatible with scikit-learn.
\subsection{Kernel quantile regression}
Kernel quantile regression had no previously implemented open source library, thus the need of implementing our own version.
\\
The scikit-learn team provides a project template for the creation of estimators compatible with scikit-learn functionalities. Therefore, the KQR class is derived from the scikit-learn BaseEstimator and the mixin class RegressorMixin.
Our KQR class is initialized by providing a quantile, the regularization term C and the rbf kernel bandwidth gamma; the latter are the two hyperparameters of our custom estimator.
\\
In the fit method, we set up and solve the convex optimisation problem by using the cvxopt library. 
When using this library, it is important to keep two things in mind. First this library assumes the quadratic term of the optimisation problem to be multiplied by the 0.5 factor, thus we just have to provide the Q matrix with no 0.5 in front.
Secondly, in order to specify multiple inequalities we have to stack them an provide as a unique matrix.
\\
Once a solution to the convex problem has been found, we create a mask for the support vectors of the estimator in order to estimate the constant term of our kernel quantile regressor.
\\
In the predict method, we pass a matrix X\_eval of independent variables, next we compute the kernel matrix between X\_train and X\_eval and obtain y\_eval with the formula $y=\alpha^\intercal K+b$.
\\
This estimator is compatible with useful scikit-learn in built methods like gridsearch, crossvalidation and scoring rules.
\\
Up to now, there are only two open source implementation of the quantile kernel regression. Nevertheless they are both in R, that is there exists no python, matlab or julia open source implementation. Following are reported the results of a comparative study between our own implementation and the one of the R library kernlab
% (i guess it is in C or C++ and then binded to R).
\subsubsection{Python vs R implementation}
In this section, a comparison study has been carried out in order to inspect the competitiveness of our implementation with the existing one for the R programming language.
